\artus -- Framework for Event-based Data Analysis in High Energy Physics

### 2015-09-10 Core/doc/latex/paper.tex
paper: some comments on the comments from RF

\begin{abstract}
We present \artus, an event-by-event data-processing framework for high energy physics experiments.
\artus is designed for large-scale data analysis in a collaborative environment.
The architecture design choices take into account typical challenges and are based on experiences with similar applications.
The structure of the framework and its advantages are described.
An example usecase and performance measurements are presented.
The framework is well-tested and successfully used by several analysis groups.\todo{Mention them, maybe with publications?}
\end{abstract}

@@

\begin{figure}[p]
\centering \input{figures/analysis_steps}
\caption[Characteristics of an event-based analysis in high energy physics]{Characteristic of an event-based analysis in high energy physics.
I/O intensive operations are done in highly parallel in the WLCG.
The computing-intensive operations are only performed if really necessary.
This means that selection steps are performed as early as possible.}
\label{figure_analysis_steps}
\end{figure}
\clearpage

@@

A pipeline consists of an arbitrary number of processor modules that are run in a sequence.
Three kinds of processors are distinguished:
\begin{description}
\item[Producers] are meant to calculate new quantities based on the information in the event and the product and following the settings of the pipeline.
\item[Filters] apply a selection based on event properties.
In case an event does not survive a requirement defined by a filter, the processing of the subsequent producers and filters can be skipped.
\item[Consumers] can be added to local pipelines.
They are executed after the producers and filters.
The task of consumers is to take quantities from the event and the product and store them in the output according to the pipeline settings.
\end{description}
Producers and filters can be configured in an arbitrary order.
Consumers are executed after the chain of producers and filters.
The output together with the complete configuration is written to a ROOT \todo{Referenz auf ROOT?} file.
These processors are created and managed by a dedicated factory according to the configuration.
\todo{I would not put the remaining part of this chapter here.
It should be either part of the requirements or part of the summary what Artus does achieve; but in this technical part in between it feels misplaced}
This structure extends the established concepts of existing frameworks like FWLite~\cite{FWLite} from CMSSW to the following four main advantages:
\begin{description}
\item[Modular structure] Analysis steps are implemented in processors, breaking down the complete analysis in well-defined and easily understandable and writeable parts.
\item[Re-utilisation of code] Processors can be shared among different analysis groups.
Every user contributes to the testing of existing code and improves the confidence in this code.
\item[Configurability] Each module uses settings that are read from the the configuration.
Each pipeline is mapped to a separate sub-set of the configuration.
\item[High performance] The concept of pipelines allows to analyse the data with different configurations at the same time, while the data is only read in only once and common processing steps are shared in the global pipeline.
Therefore the cost-intensive input operations are reduced to a minimum.
\end{description}

\section{Building an End-user Analysis \label{section_artus_analysis}}
\todo{If I wouldn't know the structure already, the following part wouldn't have helped my.
Either we support it with a drawing and more detailled explanations or we just refer to a readme...}
In the \artus framework each concrete analysis is defined by a set of object types.
It needs an implementation of the classes for settings, the event and the product.
The settings class defines the tags that should be able to be read from the configuration file and the corresponding types of the values.
The event class defines the event content being read from the input and the product class contains members for new quantities being calculated by the analysis.
These three classes need to be derived from the base classes in the core of the framework.

>>>>

\begin{abstract}
We present \artus, an event-by-event data-processing framework for high energy physics experiments.
\artus is designed for large-scale data analysis in a collaborative environment.
The architecture design choices take into account typical challenges and are based on experiences with similar applications.
The structure of the framework and its advantages are described.
An example usecase and performance measurements are presented.
The framework is well-tested and successfully used by several analysis groups.\todo{Mention them, maybe with publications? -> mentioned at the beginning of Htt chapter. Do we have publications?}
\end{abstract}

@@

\begin{figure}[p]
\centering \input{figures/analysis_steps}
\caption[Characteristics of an event-based analysis in high energy physics]{Characteristic of an event-based analysis in high energy physics.
I/O intensive operations are done in highly parallel in the WLCG.
The computing-intensive operations are only performed if really necessary.
This means that selection steps are performed as early as possible.}
\label{figure_analysis_steps}
\end{figure}
\todo{Should we drop CMS-specific technical terms like AOD/miniAOD?
Maybe not put this figure on a dedicated page?}
\clearpage

@@

A pipeline consists of an arbitrary number of processor modules that are run in a sequence.
Three kinds of processors are distinguished:
\begin{description}
\item[Producers] are meant to calculate new quantities based on the information in the event and the product and following the settings of the pipeline.
\item[Filters] apply a selection based on event properties.
In case an event does not survive a requirement defined by a filter, the processing of the subsequent producers and filters can be skipped.
\item[Consumers] can be added to local pipelines.
They are executed after the producers and filters.
The task of consumers is to take quantities from the event and the product and store them in the output according to the pipeline settings.
\end{description}
Producers and filters can be configured in an arbitrary order.
Consumers are executed after the chain of producers and filters.
The output together with the complete configuration is written to a ROOT \todo{Referenz auf ROOT?} file.
These processors are created and managed by a dedicated factory according to the configuration.

\todo{I would not put the remaining part of this chapter here.
It should be either part of the requirements or part of the summary what Artus does achieve; but in this technical part in between it feels misplaced
-> Should we put this part into a dedicated section directly after this one?
I think the idea was to explain how certain design decisions, e.g. pipelines, achieve the \artus design goals.
Therefore it shouldnt be too far from the technical descriptions.
}
This structure extends the established concepts of existing frameworks like FWLite~\cite{FWLite} from CMSSW to the following four main advantages:
\begin{description}
\item[Modular structure] Analysis steps are implemented in processors, breaking down the complete analysis in well-defined and easily understandable and writeable parts.
\item[Re-utilisation of code] Processors can be shared among different analysis groups.
Every user contributes to the testing of existing code and improves the confidence in this code.
\item[Configurability] Each module uses settings that are read from the the configuration.
Each pipeline is mapped to a separate sub-set of the configuration.
\item[High performance] The concept of pipelines allows to analyse the data with different configurations at the same time, while the data is only read in only once and common processing steps are shared in the global pipeline.
Therefore the cost-intensive input operations are reduced to a minimum.
\end{description}

\section{Building an End-user Analysis \label{section_artus_analysis}}
\todo{If I wouldn't know the structure already, the following part wouldn't have helped my.
Either we support it with a drawing and more detailled explanations or we just refer to a readme...
Maybe rewrite parts of this section}
In the \artus framework each concrete analysis is defined by a set of object types.
It needs an implementation of the classes for settings, the event and the product.
The settings class defines the tags that should be able to be read from the configuration file and the corresponding types of the values.
The event class defines the event content being read from the input and the product class contains members for new quantities being calculated by the analysis.
These three classes need to be derived from the base classes in the core of the framework.
